---
title: "post04"
author: "Mike McGrath"
date: "3/31/2020"
output:
  word_document: default
  html_document: default
---

### How have model portfolios held up after Q1 2020?

We want to look at the returns (YTD, 1Y, 5Y) for different model portfolios. We will use different model portfolios to assess the performance over the respective periods. The model portfolios are based on YCharts Basic ETF Models.

### Prepare data

First, let's set our working directory and load some libraries we might need.

```{r}

# set working directory
setwd("~/Desktop/immature_investor/post04")

# load libraries
library(dplyr)
library(tidyverse)
library(tibbletime)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)
library(quantmod)
library(PortfolioAnalytics)
library(purrr)
library(BatchGetSymbols)
library(ggrepel)

```

Now let's import and explore our data.

```{r}

# import data
all_stocks <- 
  read_excel("model_portfolios.xlsm", sheet = "100")
head(all_stocks)

eighty_twenty <- 
  read_excel("model_portfolios.xlsm", sheet = "80-20")
head(eighty_twenty)

sixty_forty <- 
  read_excel("model_portfolios.xlsm", sheet = "60-40")
head(sixty_forty)

forty_sixty <- 
  read_excel("model_portfolios.xlsm", sheet = "40-60")
head(forty_sixty)

twenty_eighty <- 
  read_excel("model_portfolios.xlsm", sheet = "20-80")
head(twenty_eighty)

all_bonds <- 
  read_excel("model_portfolios.xlsm", sheet = "100")
head(all_bonds)

```

### Calculate Returns

Now that we have imported our model portfolios, let's gather historical data returns. We can simply change the 'from' date for each period we want to look at.

All Stocks...

```{r}

# set up symbols and respective weights
all_stocks_symbols <- c("VTI", "VXUS", "0P0000MO5E.TO")
all_stocks_weights <- c(.588, .392, .02)

# find historical prices
all_stocks_prices <- NULL
for (all_stocks_symbols in all_stocks_symbols) {
all_stocks_prices = cbind(all_stocks_prices,
                        getSymbols.yahoo(all_stocks_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(all_stocks_prices)

# calculate returns
all_stocks_returns <-
  na.omit(ROC(all_stocks_prices))
head(all_stocks_returns)

# find portfolio returns
all_stocks_portfolio_returns <- 
  Return.portfolio(all_stocks_returns, weights = all_stocks_weights, rebalance_on = "years")

all_stocks_results <-
  all_stocks_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns1 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(all_stocks_results$cum_returns1)

```

80/20...

```{r}

# set up symbols and respective weights
eighty_twenty_symbols <- c("VTI", "VXUS", "BND", "BNDX", "0P0000MO5E.TO")
eighty_twenty_weights <- c(.47, .314, .137, .059, .02)

# find historical prices
eighty_twenty_prices <- NULL
for (eighty_twenty_symbols in eighty_twenty_symbols) {
eighty_twenty_prices = cbind(eighty_twenty_prices,
                        getSymbols.yahoo(eighty_twenty_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(eighty_twenty_prices)

# calculate returns
eighty_twenty_returns <-
  na.omit(ROC(eighty_twenty_prices))
head(eighty_twenty_returns)

# find portfolio returns
eighty_twenty_portfolio_returns <- 
  Return.portfolio(eighty_twenty_returns, weights = eighty_twenty_weights, rebalance_on = "years")

eighty_twenty_results <-
  eighty_twenty_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns2 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(eighty_twenty_results$cum_returns2)

# look at a quick plot of the two cumulative portfolio returns we have aready done to confirm method
all_stocks_results %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = cum_returns1), color = "red") +
  geom_line(data = eighty_twenty_results, aes(y = cum_returns2), color = "blue")


```

60/40...

```{r}

# set up symbols and respective weights
sixty_forty_symbols <- c("VTI", "BND", "VXUS", "BNDX", "0P0000MO5E.TO")
sixty_forty_weights <- c(.353, .274, .235, .118, .02)

# find historical prices
sixty_forty_prices <- NULL
for (sixty_forty_symbols in sixty_forty_symbols) {
sixty_forty_prices = cbind(sixty_forty_prices,
                        getSymbols.yahoo(sixty_forty_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(sixty_forty_prices)

# calculate returns
sixty_forty_returns <-
  na.omit(ROC(sixty_forty_prices))
head(sixty_forty_returns)

# find portfolio returns
sixty_forty_portfolio_returns <- 
  Return.portfolio(sixty_forty_returns, weights = sixty_forty_weights, rebalance_on = "years")

sixty_forty_results <-
  sixty_forty_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns3 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(sixty_forty_results$cum_returns3)

```

40/60...

```{r}

# set up symbols and respective weights
forty_sixty_symbols <- c("BND", "VTI", "BNDX", "VXUS", "0P0000MO5E.TO")
forty_sixty_weights <- c(.412, .235, .176, .157, .02)

# find historical prices
forty_sixty_prices <- NULL
for (forty_sixty_symbols in forty_sixty_symbols) {
forty_sixty_prices = cbind(forty_sixty_prices,
                        getSymbols.yahoo(forty_sixty_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(forty_sixty_prices)

# calculate returns
forty_sixty_returns <-
  na.omit(ROC(forty_sixty_prices))
head(forty_sixty_returns)

# find portfolio returns
forty_sixty_portfolio_returns <- 
  Return.portfolio(forty_sixty_returns, weights = forty_sixty_weights, rebalance_on = "years")

forty_sixty_results <-
  forty_sixty_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns4 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(forty_sixty_results$cum_returns4)

```

20/80...

```{r}

# set up symbols and respective weights
twenty_eighty_symbols <- c("BND", "BNDX", "VTI", "VXUS", "0P0000MO5E.TO")
twenty_eighty_weights <- c(.549, .235, .118, .078, .02)

# find historical prices
twenty_eighty_prices <- NULL
for (twenty_eighty_symbols in twenty_eighty_symbols) {
twenty_eighty_prices = cbind(twenty_eighty_prices,
                        getSymbols.yahoo(twenty_eighty_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(twenty_eighty_prices)

# calculate returns
twenty_eighty_returns <-
  na.omit(ROC(twenty_eighty_prices))
head(twenty_eighty_returns)

# find portfolio returns
twenty_eighty_portfolio_returns <- 
  Return.portfolio(twenty_eighty_returns, weights = twenty_eighty_weights, rebalance_on = "years")

twenty_eighty_results <-
  twenty_eighty_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns5 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(twenty_eighty_results$cum_returns5)

```

All Bonds...

```{r}

# set up symbols and respective weights
all_bonds_symbols <- c("BND", "BNDX", "0P0000MO5E.TO")
all_bonds_weights <- c(.686, .294, .02)

# find historical prices
all_bonds_prices <- NULL
for (all_bonds_symbols in all_bonds_symbols) {
all_bonds_prices = cbind(all_bonds_prices,
                        getSymbols.yahoo(all_bonds_symbols,
                        from = "2015-03-31",
                        periodicity = "daily",
                        auto.assign = FALSE) [, 6])
}
head(all_bonds_prices)

# calculate returns
all_bonds_returns <-
  na.omit(ROC(all_bonds_prices))
head(all_bonds_returns)

# find portfolio returns
all_bonds_portfolio_returns <- 
  Return.portfolio(all_bonds_returns, weights = all_bonds_weights, rebalance_on = "years")

all_bonds_results <-
  all_bonds_portfolio_returns %>%
  fortify() %>%
  na.omit() %>%
  rename(date = "Index", daily_return = "portfolio.returns") %>%
  mutate(cum_returns6 = 1 + cumsum(daily_return)) 

# visualize to check calculations
plot.ts(all_bonds_results$cum_returns6)

```

### Plot All Models

```{r}

# set ggplot theme
mike_mcgrath_theme = theme(plot.title = element_text(face = "bold", size = 15, hjust = 0.5, family = "Times New Roman"),
                           panel.background = element_rect(fill = "white"),
                           plot.caption = element_text(hjust = 0, family = "Times New Roman"),
                           axis.line = element_line(colour = "black"),
                           axis.text = element_text(colour = "black", family = "Times New Roman"),
                           axis.title = element_text(colour = "black", family = "Times New Roman"),
                           axis.ticks.length = unit(.20, "cm"),
                           panel.grid.major.y = element_line(colour = "#CCCCCC"))

# set captions
source_string = paste0("Data source: Yahoo Finance (theimmatureinvestor.wordpress.com)")
note_string = paste0("Note: Diversified portfolios are based on basic ETF models. Portfolios are rebalanced
on an annual basis. No contributions and no withdrawls are assumed.")

# set colors
color <- c("100/0" = "orchid",
           "80/20" = "gold",
           "60/40" = "steelblue",
           "40/60" = "lawngreen",
           "20/80" = "deeppink",
           "0/100" = "darkorange")

# plot
all_stocks_results %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = cum_returns1-1, color = "100/0")) +
  geom_line(data = eighty_twenty_results, aes(y = cum_returns2-1, color = "80/20")) +
  geom_line(data = sixty_forty_results, aes(y = cum_returns3-1, color = "60/40")) +
  geom_line(data = forty_sixty_results, aes(y = cum_returns4-1, color = "40/60")) +
  geom_line(data = twenty_eighty_results, aes(y = cum_returns5-1, color = "20/80")) +
  geom_line(data = all_bonds_results, aes(y = cum_returns6-1, color = "0/100")) +
  mike_mcgrath_theme +
  theme(axis.text.x = element_text(angle = 0),
        axis.ticks.x = element_blank(),
        legend.key = element_rect("white"),
        legend.title = element_blank(),
        legend.text = element_text(family = "Times New Roman"),
        legend.position = "top") +
  labs(title = "Diversified Portfolio Returns 5Y",
       x = "", 
       y = "Cumulative Return",
       caption = paste0(source_string, "\n", note_string, "\n")) +
  scale_x_date(breaks = date_breaks("years"), 
               labels = date_format("%b %Y")) +
  scale_y_continuous(labels = percent_format()) 

```





