---
title: "post03"
author: "Mike McGrath"
date: "3/28/2020"
output: html_document
---

Topic: We want to look at the drawdowns of the SP500 and discuss the importance of a financial advisor. Financial advisors earn their stripes in times of market panic. We have all heard it before, buy low sell high, or even buy and hold. But just how easy is it to simply "buy and hold" when the market is down 30%, 40%, even 50%? In this post I discuss the role of a financial advisor and how important they can be when the market isn't making new highs every week.

We will use data since the start of the most recent bull market on March 9, 2009. We are using data since the start of the bull market to highlight the fact that investors have not seen drawbacks like we are seeing now, nor the volatility. 

### Load libraries

```{r}

library(tidyverse)
library(tidyquant)
library(timetk)
library(scales)
library(highcharter)
library(broom)
library(PerformanceAnalytics)
library(dplyr)
library(tibbletime)
library(lubridate)
library(ggplot2)
library(quantmod)

```

### Import data

We will be working with SP500 (SPY).

```{r}

dd_sp500 <-
  getSymbols.yahoo("SPY",
                   from = "2009-03-09",
                   auto.assign = FALSE) [,6] 

head(dd_sp500)
```

### Drawdowns

We now have to find the drawdowns during the period. To find drawdowns we have to use quantmod. This is the easiest way known as of now. 

```{r}

dd_returns <-
  dailyReturn(dd_sp500, type = "arithmetic")

dd_returns %>%
  # get quick stats
  table.Drawdowns(top = 10)

```

### Visualize drawdowns
We now have the information we need to plot drawdowns. Let's visualize.

```{r}

# set up plot theme for graphs
mike_mcgrath_theme = theme(plot.title = element_text(face = "bold", size = 15),
                           panel.background = element_rect(fill = "white"),
                           plot.caption = element_text(hjust = 0),
                           axis.line = element_line(colour = "black"),
                           axis.text = element_text(colour = "black"),
                           axis.ticks.length = unit(.20, "cm"),
                           legend.position = "none",
                           panel.grid.major.y = element_line(colour = "#CCCCCC"))

# set up captions
source_string = paste0("Data source: Yahoo Finance (theimmatureinvestor.wordpress.com)")
note_string = paste0("Note: S&P 500 data is based on ticker symbol SPY.")

dd_returns %>%
  Drawdowns() %>%
  fortify() %>%
  rename(date = 'Index', return = 'daily.returns') %>%
  ggplot(aes(x = date)) +
  geom_ribbon(aes(ymin = return, ymax = 0),
              color = "red",
              fill = "red") +
  mike_mcgrath_theme +
  theme(axis.text.x = element_text(angle = 45),
        axis.ticks.x = element_blank()) + 
  labs(title = "S&P 500 Drawdowns Since 2009",
       x = "",
       y = "Drawdown",
       caption = paste0(source_string, "\n", note_string, "\n")) +
  scale_y_continuous(labels = percent_format(accuracy = 2)) +
  scale_x_date(breaks = date_breaks("years"),
               labels = date_format("%Y"))
  
```

Let's plot the SP500 daily returns. We are looking to show the volatility we have seen this year and how it is unlike anything we have seen in recent years.

We also want to create an 'if_else()' statement that does the following:
*If* daily return is >= +/-4%, 1, *else* 0 

Let's set up the 'if_else()' statement first. 

```{r}
dd_returns %>%
  fortify() %>%
  rename(date = 'Index', return = 'daily.returns') %>% 
  # add our logic
  mutate(color = if_else(abs(return) >= .04, "red", "black")) 

```

Now let's plot.

```{r}

# set up captions
source_string = paste0("Data source: Yahoo Finance (theimmatureinvestor.wordpress.com)")
note_string = paste0("Note: S&P 500 data is based on ticker symbol SPY. Red dots correspond to 
daily returns of (+/-) 4.0%.")

dd_returns %>%
  fortify() %>%
  rename(date = 'Index', return = 'daily.returns') %>%
  # add our logic
  mutate(color = if_else(abs(return) >= .05, "red", "black")) %>%
  ggplot(aes(x = date)) +
  geom_point(aes(y = return, col = color)) +
  scale_colour_identity() +
  labs(title = "S&P 500 Daily Returns Since 2009",
       x = "",
       y = "Daily Return",
              caption = paste0(source_string, "\n", note_string, "\n")) +
  mike_mcgrath_theme +
  theme(axis.text.x = element_text(angle = 45),
        axis.ticks.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 2)) +
  scale_x_date(breaks = date_breaks("years"),
               labels = date_format("%Y"))

```





